Diagnostic Plots

### 2024-07-23 response_to_reviewers.qmd

---
title: "Response to Reviewers"
subtitle: "Updating meta-analyses cumulatively using Bayesian methods: An example updating a previous meta-analysis of self-talks interventions for sport/motor performance"
format: pdf
---

Dear \[insert editor name\],

Many thanks for considering our manuscript and placing it under review with \[insert journal name\]. We are very grateful for your efforts in obtaining reviews. We are very impressed with the quality and constructiveness of the feedback provided by the reviewers which we feel has significantly improved the quality of our manuscript.

Note, because of the extensive rewrite of the manuscript due to the shift in focus we have not used track changes or highlighting throughout, though changes can be viewed in the GitHub repository versions. However, please see below for our comments in response to some of the key points raised by the reviewers.

Again, we would like to thank all involved in the process of reviewing this manuscript and look forward to the next round of reviews in considering our revisions.

Many thanks

The authors

# Reviewer Comments and [Author Responses]{style="color:red;"}

## Reviewer 1 - Maik Bieleke

### Reviewer Comment

Contrary to what the title suggests, I did not have the impression that the article is best framed as a meta-analysis of self-talk interventions. It does not align particularly well with established standards for meta-analyses (PRISMA guidelines; e.g. flow charts, risk-of-bias analyses) and there are only very limited references to the self-talk literature. Rather, the article seems better described as a methodological contribution that demonstrates the benefits of Bayesian methods for cumulative science, using self-talk interventions as a convenient but largely interchangeable example. The keywords chosen by the authors reinforce this impression by omitting any reference to self-talk and meta-analysis. Therefore, a clearer and more consistent positioning of the article as valuable methodological contribution to cumulative science, rather than a standalone meta-analysis, might enhance clarity in this regard.

### [Author Response]{style="color:red;"}

[Thank you for your suggestion. It is actually a point which in retrospect after submitting the manuscript we realised more ourselves and so it is gratifying to have this reinforced by both reviewers. For clarity, the absence of certain methodological details such as those noted did stem partly from us wanting to replicate the methods employed by Hatzigeorgiadis et al. as closely as possible. As this was an undergraduate project being completed in a short period of time we opted to try to replicate to make the project more efficient under such constraints. We did initially intend though to include a PRISMA flow diagram to describe our updated searches, however we encountered an issue and realised that some of the searches we had conducted were not correctly recorded by the tool being used to manage the process. Unfortunately we were unable to exactly replicate it from attempting to reproduce the searches again. We have added this detail to footnote 2 for transparency.]{style="color:red;"}

[We do agree though that positioning this paper instead as an example of using Bayesian meta-analytic methods for cumulative science, whilst using the self-talk literature as an exemplar case-study given how long it has been since a quantitative evidence synthesis was conducted, is perhaps more suitable. As such we have attempted to rewrite the manuscript with this in mind, and have also changed the title to: *Cumulative evidence synthesis using Bayesian methods: An example updating a previous meta-analysis of self-talks interventions for sport/motor performance*]{style="color:red;"}

### Reviewer Comment

Whether or not the authors follow my previous suggestion, I think that a brief introduction to research on self-talk and self-talk interventions would aid readers in comprehending the narrative of the article (e.g., understanding and interpreting the moderator analyses). For example, the term self-talk is never explicitly introduced or defined, and the "key distinction" (p. 13) between strategic and organic self-talk is only mentioned towards the end of the discussion.

### [Author Response]{style="color:red;"}

[Given the reframing of the paper, we have now subsectioned the introduction and include a section more explicitly introducing self-talk.]{style="color:red;"}

### Reviewer Comment

On a quantitative note, I found it surprising that the confidence interval around the overall effect size remained virtually unchanged, although the amount of evidence doubled in terms of studies and effect sizes. Interestingly, the posterior CIs of the various moderators were indeed narrower than the prior CIs, an increase in precision one would expect when sample sizes increase. I wondered whether this might reflect unobserved qualitative differences between studies of some kind (e.g., additional moderators; also see comment 6 below) but maybe there is a simpler explanation that I'm missing.

### [Author Response]{style="color:red;"}

[This is an interesting, and somewhat counterintuitive result of our analysis. From the additional work we have conducted whereby we have utilised simulation to add a single new study and explore it's effects upon the prior distribution we were surprised that even a study simulated with one millions participants and a null effect had next to no impact in updating the prior estimate. As a result we conducted a little additional exploration regarding how the prior distribution was set. We utilised a $t$-distribution ($t(k,\mu,\sigma)$) with $k-2$ degrees of freedom as suggested by Higgins et al. (2009) cited in our manuscript. But, given the large degrees of freedom due to the large number of studies in the previous meta analysis from Hatzigeorgiadis et al., the $t$-distribution approximates a normal distribution more closely and thus has tight tails which influence the impact that an additional evidence might have on the prior when updating to a posterior (see this example from Richard McElreath - <https://x.com/rlmcelreath/status/1701165075493470644?s=20>). Out of curiosity we simulated a single study of one million participants and a null effect with a more skeptical $t$-distribution using only 3 degrees of freedom thus allowing the tails to be fatter. Even with this though \~85% of the entire posterior distribution was still within the 95% interval for the prior. To our interpretation this goes to show how certain the estimate from Hatzigeorgiadis et al. already was and highlights that further intervention research could be considered wasteful (assuming it has not contributed anything else to our understanding). However, as we explain in response to reviewer 2, we have now examined using various methods the possibility of small sample and publication bias in the studies included since 2011. This paints a concerning picture to our interpretation, though one that is perhaps not that surprising given that there is widespread evidence of such questionable research practices across many fields including the sport and exercise sciences. Hatzigeorgiadis et al. only examined this possibility using the now known to be flawed fail-safe N and from it concluded that there was likely no publication bias present prior to 2011. We find this hard to believe given that it is highly likely that prior to 2011, when many consider the current replication crisis in psychology (of course which sport psychology sits close to) kicked off, there were not many of the methodological reforms such as pre-registration and even registered reports, open data, materials, and code etc. in place that might help to mitigate somewhat the influence of QRPs such as publication bias. We didn't really want to end up down this route with the paper, and as noted due to it being an undergraduate project had deliberately conducted the updated systematic search to reduce the burden given the time available to complete the work. So we have not replicated the systematic search of Hatzigeorgiadis et al. for pre-2011. However, we have contacted the authors and asked for whether they still have their extracted data such that we might be able to examine this in the pre-2011 studies.]{style="color:red;"}

```         
    ### Reviewer Comment

    In the moderators section, the authors assert minimal differences between posterior estimates and priors. However, looking at Figure 2 it seems that the most extreme differences consistently decreased across moderator variables, pushing the posterior distributions closer to the overall effect size. Admittedly, this visual impression may be challenging to quantify or test, but it would be interesting to know the authors interpretation.

    ### [Author Response]{style="color:red;"}
    []{style="color:red;"}
```

### Reviewer Comment

I wondered whether "research waste" really is an appropriate label for all of the studies included in the present analysis. While these studies might have replicated the self-talk intervention effect, maybe this was not their main and/or only goal? Isn't it possible that these studies relied on the already established effect to answer a different question? For example, one might think of studies comparing self-talk as established intervention to a novel intervention. Such a study would likely be included in the present analysis due to its replication of the self-talk intervention effect, but its contribution mainly relates to examining the novel intervention. The term "research waste" then might overlook important contributions from these studies that go beyond (possibly unnecessary) replication efforts.

Relatedly, it might be worth noting that the analysis focused solely on quantitative aspects of replication (effect sizes, confidence intervals). This is fine but the authors also mention theoretical and conceptual advancements in the literature on self-talk. For somebody unfamiliar with the self-talk literature, it seems implausible that such improvements did not affect in any way how post-Hatzigeorgiadis self-talk interventions were delivered and/or evaluated. However, if administration or evaluation methods have changed (and potentially improved), the fact that our beliefs about intervention effects still remain mostly unchanged actually is reassuring, no? A qualitative analysis of the included studies might go beyond the scope of the present study, but it is still worth discussing whether and how it limits the interpretation of the quantitative analysis.

### [Author Response]{style="color:red;"}

[This is certainly a possibility and we have tried to soften the language regarding "research waste" somewhat throughout and also emphasise this limitation more in the discussion. We had considered in light of this comment whether to conduct a qualitative appraisal of the included studies to determine the extent to which they had either merely estimated the effects of self-talk interventions or explored moderators which were already well understood (though this we now question due to the possibility of publication bias in main effects) as opposed to contributing something new theoretically or methodologically. However, given the shift in the focus of the manuscript we opted not to do so but instead to highlight this as something those engaged in cumulative evidence synthesis should consider as well when determining whether research could be considered wasteful, or in the justification of a new trial by dint of the fact it might address theoretically or methodologically novel considerations.]{style="color:red;"}

### Reviewer Comment

In the introduction, the authors argue that "the effect estimate from the meta-analysis of Hatzigeorgiadis et al. (2011) was already fairly precise", implying that further research was unwarranted to begin with. I found the statement rather vague and the implication drawn with the benefit of hindsight. When is an estimate precise enough to discourage further replications, and aren't there other reasons for replicating an effect (e.g., robustness checks, changes in methodology) that warrant investigation?

### [Author Response]{style="color:red;"}

[We have now attempted to address this concern through the addition of simulations of the effects that a new study might have had post the meta-analysis of Hatzigeorgiadis et al. in terms of changing our conclusions regarding the effect. This lends some quantitative precision the the claim that it was already fairly precise adn that further research may have been unwarranted. Of course there may be other reasons to continue further research attempting to replicate an effect, though in this case we'd be skeptical that it is warranted.]{style="color:red;"}

### Reviewer Comment

There is currently no discussion of the potential strengths and limitations of the study. I think that adding this would make it easier for readers to gauge the contribution of the present research.

### [Author Response]{style="color:red;"}

[Given the reframing we instead have added discussion of the benefits of engaging with the kinds of methods demonstrated, but have also noted that supplementing this with qualitative appraisal is of value in determining wwhether addtional research is warranted.]{style="color:red;"}

## Reviewer 2 - Anonymous

### Reviewer Comment

Meta-analyses (as the authors of course know) estimate an average effect and deviations from this effect. Thus, meta-analyses only make sense when it is theoretically or conceptually meaningful to estimate an average effect over a set of studies. The Data Colada team makes this point better than I could, and they illustrate it with some nice examples (please see http://datacolada.org/104 and the related posts).

So, I was wondering: Does it make sense (both from a theoretical and an applied perspective) to estimate the average effect of self-talk? I am not saying that it doesn't, I simply do not know. In other words: The more the different self-talk interventions resemble each other (both in terms of theoretical foundations and in terms of their implementation), the more sense it makes to conduct a meta-analysis regarding their average effectiveness. Maybe the authors can address this point and add a couple of words on the similarity of the analyzed interventions.

### [Author Response]{style="color:red;"}

[This is a good question and one which we encounter often. Specifically in a random effects meta-analysis the overall main estimate is the weighted average of the effect*s* that each study has estimated. It begins with the assumption that the specific effect being estimated does differ across the studies for a range of factors mostly relating to the fact that they are all conceptual replications as opposed to direct replications. The benefit though is that we can quantify the heterogeneity via the $\tau$ parameter and determine whether it is meaningful enough to be concerned about. In addition, the moderator analyses allow for us to explore whether or not study, experiment, or group level factors might be influences of the effects that each study estimates within the overarching distribution of effects. So, we would argue that in the case of asking the question about the effects of an intervention, such as self-talk upon an outcome, such as sport/motor performance, meta-analysis can be appropriate if conducted appropriately.]{style="color:red;"}

### Reviewer Comment

I was a bit surprised that the authors did not assess potential publication bias, but maybe I missed something. It is my understanding that publication bias (e.g., only positive results are published) poses a serious threat to meta-analyses, and that therefore tools to assess publication bias have been developed. These tools may range from funnel plots to techniques such as Trim-and-fill, PET, PEESE and so on (e.g., Carter, E. C., Schönbrodt, F. D., Gervais, W. M., & Hilgard, J. (2019). Correcting for bias in psychology: A comparison of meta-analytic methods. Advances in Methods and Practices in Psychological Science, 2, 115-144. https://doi.org/10.1177/2515245919847196). Why did the authors not address potential publication bias? As far as I am concerned, one reason why results have not changed from 2011 till now might be that results that fit the picture get published easier?

### [Author Response]{style="color:red;"}

[Thank you for this suggestion. Somewhat in our haste and our focus on the updating aspect of the novel approach taken, we failed to consider this. However, we have now examined using various methods the possibility of small sample and publication bias in the studies included since 2011. The PET-PEESE and Robust Bayesian Model Averaging methods reveal evidence of publication bias and that the adjusted effect may in fact be null. This paints a concerning picture to our interpretation, though one that is perhaps not that surprising given that there is widespread evidence of such questionable research practices across many fields including the sport and exercise sciences. Hatzigeorgiadis et al. only examined this possibility using the now known to be flawed fail-safe N and from it concluded that there was likely no publication bias present prior to 2011. We find this hard to believe given that it is highly likely that prior to 2011, when many consider the current replication crisis in psychology (of course which sport psychology sits close to) kicked off, there were not many of the methodological reforms such as pre-registration and even registered reports, open data, materials, and code etc. in place that might help to mitigate somewhat the influence of QRPs such as publication bias. We didn't really want to end up down this route with the paper, and as noted due to it being an undergraduate project had deliberately conducted the updated systematic search to reduce the burden given the time available to complete the work. So we have not replicated the systematic search of Hatzigeorgiadis et al. for pre-2011. However, we have contacted the authors and asked for whether they still have their extracted data such that we might be able to examine this in the pre-2011 studies.]{style="color:red;"}

### Reviewer Comment

I do not understand the authors' point that all research in the meantime between the first meta-analysis and theirs was "a waste of research", simply because their meta-analysis comes to the same conclusion as the first one. To me, this statement is somewhat questionable for several reasons (at least in the way it is presented in the present form of the article).

First, the argument is only logically valid if we suppose that the *only* goal of the studies that were analyzed was to assess the effectiveness of self-talk. However, if some of the studies had additional goals, then the "waste of research" conclusion is not valid anymore. Please let me illustrate this point with a couple of examples. Suppose that a study aimed to find out whether self-talk works better for soccer players than for basketball players. The results show that it doesn't. In this case, we have learned something from the respective study, even if in a meta-analysis the study gets the same effect size like in a previous meta-analysis.

Or suppose that a study aimed to find out whether self-talk works better for handball players than for volleyball players. The results suggest that it works above average for handball players and below average for volleyball players. Again, the average effect would be the same as in the previous meta-analysis, but again I assume that we have learned something.

On a very basic level, I would like to argue that in order to decide whether a study was a waste of research and resources you have to take into account that very study's goals. Therefore, I was wondering: Was the only goal of the studies analyzed in the present meta-analysis really to assess the average effectiveness of self-talk?

### [Author Response]{style="color:red;"}

[Reviewer 1 also raised this concern and we have tried to soften the language regarding "research waste" somewhat throughout and also emphasise this limitation more in the discussion. We had considered in light of this comment whether to conduct a qualitative appraisal of the included studies to determine the extent to which they had either merely estimated the effects of self-talk interventions or explored moderators which were already well understood (though this we now question due to the possibility of publication bias in main effects) as opposed to contributing something new theoretically or methodologically. However, given the shift in the focus of the manuscript we opted not to do so but instead to highlight this as something those engaged in cumulative evidence synthesis should consider as well when determining whether research could be considered wasteful, or in the justification of a new trial by dint of the fact it might address theoretically or methodologically novel considerations.]{style="color:red;"}

### Reviewer Comment

When exactly does additional research become "a waste"? I totally agree that research is not efficient anymore when it does not add anything to established knowledge anymore. However, when is knowledge really established in psychology? Considering the average sample sizes in psychology, even 10 psychological studies might not have the sample size that might be considered necessary to yield established knowledge in other fields. Again, I am not saying that there will not eventually come a point where further research becomes unnecessary and thus inefficient. I am simply wondering: When does it come?

### [Author Response]{style="color:red;"}

[We have tried to better address this by leveraging simulating the addition a single new study, varying the underlying effect size and sample size, and explore it's effects upon the prior distribution. This is to demonstrate that such an approach can aid in determining exactly when the addition of new studies, whether small or indeed large, might be unnecessary in light of current evidence.]{style="color:red;"}

### Reviewer Comment

Third, the authors themselves write (p. 2 / 3) that research on self-talk might have matured post 2011 regarding "... efforts to improve operationalisation/measurement, and efforts to improve methodology used in studying self-talk". Quite frankly, I cannot reconcile this observation with the statement that this research was a waste. I mean, finding the same effects with better measurement and methods should increase our confidence both in self-talk per se, but also in the results prior to 2011, or not? Suppose you have a study that finds a certain effect, but this study has weaknesses. Now a study with better methods finds the same effect. Would you really want to argue that the second study was a waste, because it finds the same effect? I surely would not.

### [Author Response]{style="color:red;"}

[As noted we have tried to make it clear that such qualitative considerations should also be taken into account when determining whether research has been wasteful, or indeed further research might be. Although, for the specific example of self-talk the process of working on this project has been enlightening and left us in considerable doubt of the value of much of the research in the area (though notably that would be the case for many areas within the sport and exercise sciences).]{style="color:red;"}

### 2024-07-23 pre_print_manuscript.qmd
Fix the dang "onto to" abstract typo.
Rebuilding targets pipeline... up to motor demands up to date

---
articletitle: |
  Self-talk interventions and sport/motor performance: An updated systematic review and Bayesian meta-analysis
format: 
  sportrxiv-pdf:
    include-in-header:
      text: |
       \usepackage[font=scriptsize]{caption}
       \usepackage{lscape}
        \newcommand{\blandscape}{\begin{landscape}}
        \newcommand{\elandscape}{\end{landscape}}
author:
  - name: Hannah Corcoran
    affiliations:
      - ref: 1
  - name: James Steele
    affiliations:
      - ref: 1
    orcid: 0000-0002-8003-0757
    corresponding: true
    email: james.steele@solent.ac.uk
affiliations:
      - id: 1
        name: Department of Sport and Health, Solent University, UK
abstract: |
  Self-talk has been researched as an aspect of mental preparation for performance in sports and other motor tasks since the late 1980s. In 2011, Hatzigeorgiadis and colleagues systematically reviewed and meta-analysed three decades of self-talk intervention research including 32 studies. Yet, despite the general proliferation of meta-analyses, this topic has not been meta-analysed in the decade since their review which, at least for the main effect, provided a reasonably precise standardised mean difference estimate (0.48 [95% confidence interval: 0.38, 0.58]). Several further studies on self-talk interventions have been conducted in that time and it is of interest to explore what additional evidence they offer regarding the effects of self-talk interventions on sport/motor performance. Bayesian approaches are well positioned to explore how additional evidence changes our understanding of an effect; to see whether it has changed in sign, magnitude, or precision, or whether further research has largely been a ‘waste’. Therefore, our aim was to conduct an updated systematic review and Bayesian meta-analysis replicating the search, inclusion, and models of Hatzigeorgiadis et al. (2011). Informative priors were taken directly from Hatzigeorgiadis et al. A total of 34 studies providing 128 effects nested in 64 groups across experiments 42 were included in the final updated meta-analysis representing data from 18761 participants. The overall posterior pooled estimate for the standardised mean difference was almost exactly the same as the prior: 0.47 [95% quantile interval: 0.39, 0.56]. Bayes factors were calculated for a range of effect sizes and indicated that the included studies largely reflected ‘Weak’ evidence against effects ranging from 0.30 to 0.59, and only provided ‘Decisive’ evidence or greater against more extreme effects: either very small (i.e., <0.02) or large (i.e., >0.81). Results were largely similar for all moderator analyses which were updated too; either providing relatively weak evidence against effects found in the previous meta-analysis or evidence suggesting smaller effects for certain moderators. The findings of our updated Bayesian meta-analyses reiterate the positive effect of self-talk interventions on sport/motor performance. However, they also suggest that cumulatively the past decade and more of research has done little to further our understanding of these effects. Considering the limited resources and time for conducting research, it may be worth moving onto to other more pertinent questions regarding psychological constructs impact upon sport/motor performance.
license-type: ccby # change if neccessary
# year: 2025 # defaults to current year
keywords: [psychological intervention, cumulative evidence, research waste] # optional
# optional link to Supplementary Material:
suppl-link: https://osf.io/dqwh5/ 
reference-section-title: References
printnote: "PREPRINT - NOT PEER REVIEWED" # if post print, include "POSTPRINT" then link to the published article
bibliography: references.bib  
pdf-engine: xelatex
execute: 
  echo: false
  message: false
  warning: false
---

@@

# Introduction

Self-talk has a long history of philosophical, theoretical, and empirical work [@geurtsMakingSenseSelf2018; @brinthauptSelftalkResearchChallenges2023; @latinjakSelfTalkInterdisciplinaryReview2023].
One area in which self-talk remains a popular topic of research up to this day is in the sport sciences [@vanraalteSelftalkReviewSportspecific2016; @hardyReflectionsMaturingResearch2018; @latinjakSelfTalkInterdisciplinaryReview2023].
Sport psychology as a broad field has focused on the theorising of psychological constructs that might impact upon performance, and the subsequent experimental testing of theoretically informed interventions to address these constructs.
For example, a recent umbrella review identified thirty meta-analyses exploring the effects of different sport psychology constructs upon performance, thirteen of them examining the effects of interventions, finding an overall standardised mean difference (SMD) for positive constructs of 0.51 \[95% confidence interval: 0.42, 0.58\] [@lochbaumSportPsychologyPerformance2022].
Only one meta-analysis explored the effects of self-talk interventions; Hatzigeorgiadis et al. [-@hatzigeorgiadisSelfTalkSportsPerformance2011].

The meta-analysis by Hatzigeorgiadis et al. [-@hatzigeorgiadisSelfTalkSportsPerformance2011] included a total of 32 studies and 62 effect size estimates revealing an overall SMD estimate of 0.48 \[95% confidence interval: 0.38, 0.58\] and also explored the effects of various moderators of the effectiveness of self-talk interventions.
Around the time that Hatzigeorgiadis et al. conducted their meta-analysis the quantitative synthesis of research findings using meta-analytic tools was still relatively new in the sport sciences [@haggerMetaanalysisSportExercise2006].
However in the last decade, particularly in sport psychology, there has been an increasing reliance on meta-analyses [@lochbaumSportPsychologyPerformance2022; @haggerMetaanalysis2022].
Yet, despite the general proliferation of meta-analyses in the past decade, the effects of self-talk interventions has not been re-evaluated by means of such quantitative synthesis since 2011, when Hatzigeorgiadis et al. [-@hatzigeorgiadisSelfTalkSportsPerformance2011] completed their work.
During this period though, empirical research regarding self-talk interventions for sport and motor performance has burgeoned leading some to reflect on the field as "maturing" post-2011 [@hardyReflectionsMaturingResearch2018].

Whilst self-talk as a field may have matured in the post-2011 years with theoretical advancements in conceptualisation of the construct and mediators of its effects on performance, efforts to improve operationalisation/measurement, and efforts to improve methodology used in studying self-talk [@brinthauptSelftalkResearchChallenges2023; @geurtsMakingSenseSelf2018; @hardyReflectionsMaturingResearch2018; @latinjakSelfTalkInterdisciplinaryReview2023; @vanraalteSelftalkReviewSportspecific2016; @latinjakSpeakingClearly102019], it could be argued that understanding of the effectiveness of self-talk *interventions* (referred to in modern literature as 'strategic' self-talk; @latinjakSpeakingClearly102019) was "mature" prior to 2011.
The effect estimate from the meta-analysis of Hatzigeorgiadis et al. [-@hatzigeorgiadisSelfTalkSportsPerformance2011] was already fairly precise, and indeed so to were many of the moderator estimates.
Several further studies on self-talk interventions have been conducted since 2011 and it is a reasonable question to ask, given the limited time and resource for conducting research in the field of sport science, to whether and to what extent these have advanced our understanding or whether they have largely contributed to "research waste".

Two questions that should (though arguably are not often enough particularly in sport science) be asked by researchers when planning a study of an experimental intervention is "what is the likelihood that the experimental intervention is superior to the control intervention given the evidence accumulated so far?", and "what is the likelihood that a new trial, given some design parameters and previous evidence, will demonstrate the superiority of the experimental intervention?".
The key here is to consider the *cumulative* nature of evidence provided by research.
Cumulative meta-analyses were proposed in the early 1990s and have since then been argued to be key tools to answering these questions and considering whether or not additional research is a worthwhile use of resources [@clarkeAccumulatingResearchSystematic2014; @graingerEvidenceSynthesisTackling2020]. Bayesian approaches are well positioned to tackle this [@biauUsingBayesianStatistics2017].
Within Bayesian statistical inference a prior probability distribution regarding the effect of interest is *updated* after the introduction of new evidence to a posterior probability distribution given Bayes theorem.

Given that there has not been, to the best of our knowledge, a meta-analytic synthesis of the effects of self-talk interventions effects upon sport/motor performance since Hatzigeorgiadis et al. [-@hatzigeorgiadisSelfTalkSportsPerformance2011] it is unclear the extent to which the past decade and more of research has advanced our understanding or potentially contributed to “research waste”.
Therefore, our aim was to conduct an updated systematic review and Bayesian meta-analysis replicating the search, inclusion, and models of Hatzigeorgiadis et al. [-@hatzigeorgiadisSelfTalkSportsPerformance2011].

>>>>

---
articletitle: |
  Cumulative evidence synthesis using Bayesian methods: An example updating a previous meta-analysis of self-talk interventions for sport/motor performance
format: 
  sportrxiv-pdf:
    include-in-header:
      text: |
       \usepackage[font=scriptsize]{caption}
       \usepackage{lscape}
        \newcommand{\blandscape}{\begin{landscape}}
        \newcommand{\elandscape}{\end{landscape}}
author:
  - name: Hannah Corcoran
    affiliations:
      - ref: 1
  - name: James Steele
    affiliations:
      - ref: 1
    orcid: 0000-0002-8003-0757
    corresponding: true
    email: james.steele@solent.ac.uk
affiliations:
      - id: 1
        name: Department of Sport and Health, Solent University, UK
abstract: |
  Self-talk has been researched as an aspect of mental preparation for performance in sports and other motor tasks since the late 1980s. In 2011, Hatzigeorgiadis and colleagues systematically reviewed and meta-analysed three decades of self-talk intervention research including 32 studies. Yet, despite the general proliferation of meta-analyses, this topic has not been meta-analysed in the decade since their review which, at least for the main effect, provided a reasonably precise standardised mean difference estimate (0.48 [95% confidence interval: 0.38, 0.58]). Several further studies on self-talk interventions have been conducted in that time and it is of interest to explore what additional evidence they offer regarding the effects of self-talk interventions on sport/motor performance. Bayesian approaches are well positioned to explore how additional evidence changes our understanding of an effect; to see whether it has changed in sign, magnitude, or precision, or whether further research has largely been a ‘waste’. Therefore, our aim was to conduct an updated systematic review and Bayesian meta-analysis replicating the search, inclusion, and models of Hatzigeorgiadis et al. (2011). Informative priors were taken directly from Hatzigeorgiadis et al. A total of 34 studies providing 128 effects nested in 64 groups across experiments 42 were included in the final updated meta-analysis representing data from 18761 participants. The overall posterior pooled estimate for the standardised mean difference was almost exactly the same as the prior: 0.47 [95% quantile interval: 0.39, 0.56]. Bayes factors were calculated for a range of effect sizes and indicated that the included studies largely reflected ‘Weak’ evidence against effects ranging from 0.30 to 0.59, and only provided ‘Decisive’ evidence or greater against more extreme effects: either very small (i.e., <0.02) or large (i.e., >0.81). Results were largely similar for all moderator analyses which were updated too; either providing relatively weak evidence against effects found in the previous meta-analysis or evidence suggesting smaller effects for certain moderators. The findings of our updated Bayesian meta-analyses reiterate the positive effect of self-talk interventions on sport/motor performance. However, they also suggest that cumulatively the past decade and more of research has done little to further our understanding of these effects. Considering the limited resources and time for conducting research, it may be worth moving onto other more pertinent questions regarding psychological constructs impact upon sport/motor performance.
license-type: ccby # change if neccessary
# year: 2025 # defaults to current year
keywords: [psychological intervention, cumulative evidence, research waste] # optional
# optional link to Supplementary Material:
suppl-link: https://osf.io/dqwh5/ 
reference-section-title: References
printnote: "PREPRINT - NOT PEER REVIEWED" # if post print, include "POSTPRINT" then link to the published article
bibliography: references.bib  
pdf-engine: xelatex
execute: 
  echo: false
  message: false
  warning: false
---

@@

###### STUFF TO ADD??
https://systematicreviewsjournal.biomedcentral.com/articles/10.1186/s13643-022-02096-y

https://www.thelancet.com/journals/lancet/article/PIIS0140-6736(15)00489-4/fulltext

https://www.thelancet.com/journals/lancet/article/PIIS0140-6736(15)01354-9/fulltext

https://academic.oup.com/biometrics/article/79/1/319/7478096?login=false

https://www.researchgate.net/publication/338878953_p-Hacking_and_Publication_Bias_Interact_to_Distort_Meta-Analytic_Effect_Size_Estimates

https://www.researchgate.net/publication/336963847_Toward_a_common_language_categorization_and_better_assessment_in_self-talk_research_Commentary_on_Speaking_clearly_10_years_on

https://www.researchgate.net/publication/312934559_On_Investigating_Self-Talk_A_Descriptive_Experience_Sampling_Study_of_Inner_Experience_During_Golf_Performance

https://www.researchgate.net/publication/337822503_Strengthening_the_Assessment_of_Self-Talk_in_Sports_Through_a_Multi-Method_Approach

https://www.researchgate.net/publication/304398817_Why_Self-Talk_Is_Effective_Perspectives_on_Self-Talk_Mechanisms_in_Sport

# Introduction
## Cumulative evidence synthesis 
Two questions that should be asked (though arguably are not often enough particularly in sport science) by researchers when planning a study of an experimental intervention is "what is the likelihood that the experimental intervention is superior to the control intervention given the evidence accumulated so far?", and "what is the likelihood that a new trial, given some design parameters and previous evidence, will demonstrate the superiority of the experimental intervention?".
The key here is to consider the *cumulative* nature of evidence provided by research and its synthesis. Indeed, to not do this could lead to so called "research waste". Evidence synthesis methods are essential to determining whether or not there is justification for further research on a given topic, and the Cochrane-Collaboration and REWARD (Reduce Research Waste and Reward Diligence) Alliance have even established an award for efforts in the area of reducing "research waste" [@glasziouResearchWasteStill2018].

          Add about QRPs and research waste - evaluate the not only the results, but the quality of the preceding evidence in addition to whether it may have been subject to any bias

Cumulative meta-analyses were proposed in the early 1990s and have since then been promoted as key tools to understand whether or not additional research is a worthwhile use of resources for addressing a particular question regarding experimental intervention [@clarkeAccumulatingResearchSystematic2014; @graingerEvidenceSynthesisTackling2020]. Further, Bayesian approaches are well positioned to tackle this [@biauUsingBayesianStatistics2017].

Within Bayesian statistical inference a prior probability distribution regarding the effect of interest is *updated* after the introduction of new evidence to a posterior probability distribution given Bayes theorem. In the present paper we demonstrate the application of Bayesian methods to meta-analysis in the sport and exercise sciences for the evaluation of experimental interventions using cumulative evidence synthesis. We assume some prior knowledge of evidence synthesis and meta-analytic methods on the part of the reader, though for those unfamiliar suggest some recent introductory papers regarding their application in the field [@gunnellQuestionsAnswersConducting2020; @haggerMetaanalysis2022; @steeleMetaanalysisVariationSport2023]. We use the example of self-talk interventions and their effects upon motor/sport performance given a quantitative evidence synthesis has not been conducted on this topic to the best of our knowledge since 2011 when Hatzigeorgiadis et al. [-@hatzigeorgiadisSelfTalkSportsPerformance2011] conducted their systematic review and meta-analysis.

## Self-talk interventions
Sport psychology as a broad field has focused on the theorising of psychological constructs that might impact upon performance, and the subsequent experimental testing of theoretically informed interventions to address these constructs.
For example, a recent umbrella review identified thirty meta-analyses exploring the effects of different sport psychology constructs upon performance, thirteen of them examining the effects of interventions, finding an overall standardised mean difference (SMD) for positive constructs of 0.51 \[95% confidence interval: 0.42, 0.58\] [@lochbaumSportPsychologyPerformance2022]. One construct with a long history of philosophical, theoretical, and empirical work [@geurtsMakingSenseSelf2018; @brinthauptSelftalkResearchChallenges2023; @latinjakSelfTalkInterdisciplinaryReview2023] that has been the target of considerable investigation in this field has been self-talk.

As a concept self-talk has been  defined in various ways in previous work on the topic; though recent transdisciplinary review [@latinjakSelfTalkInterdisciplinaryReview2023] has agreed upon a broad conceptualisation: *"verbalizations addressed to the self, overtly or covertly, characterized by interpretative elements associated to their content; and it [self-talk] either (a) reflects dynamic interplays between organic, spontaneous and goal-directed cognitive processes or (b) conveys messages to activate responses through the use of predetermined cues developed strategically, to achieve performance-related outcomes"* [@latinjakSpeakingClearly102019]. However as noted, whilst there has been various narrative syntheses of research on self-talk [@vanraalteSelftalkReviewSportspecific2016; @hardyReflectionsMaturingResearch2018; @latinjakSelfTalkInterdisciplinaryReview2023], only one systematic review and meta-analysis has explored the effects of self-talk interventions; Hatzigeorgiadis et al. [-@hatzigeorgiadisSelfTalkSportsPerformance2011].

The meta-analysis by Hatzigeorgiadis et al. [-@hatzigeorgiadisSelfTalkSportsPerformance2011] included a total of 32 studies and 62 effect size estimates revealing an overall SMD estimate of 0.48 \[95% confidence interval: 0.38, 0.58\] and also explored the effects of various theoretically driven moderators of the effectiveness of self-talk interventions. For example; characteristics of the tasks performed such as their novelty and whether they are fine or gross motor task, characteristics of the participants such as their level of experience with the task, the characteristics of the self-talk used including its content, whether it was self-selected or assigned, and whether it was used overtly or not, characteristics of the intervention and whether it included brief exposure or a training period, and testing the 'matching hypothesis' which posits that instructional self-talk should benefit fine tasks whereas motivational self-talk should benefit gross tasks to a greater degree.

Around the time that Hatzigeorgiadis et al. conducted their meta-analysis the quantitative synthesis of research findings using meta-analytic tools was still relatively new in the sport sciences [@haggerMetaanalysisSportExercise2006].
However in the last decade, particularly in sport psychology, there has been an increasing reliance on meta-analyses [@lochbaumSportPsychologyPerformance2022; @haggerMetaanalysis2022].
Despite the general proliferation of meta-analyses in the past decade, the effect of self-talk interventions has not been re-evaluated by means of such quantitative synthesis since 2011, when Hatzigeorgiadis et al. [-@hatzigeorgiadisSelfTalkSportsPerformance2011] completed their work.
During this period though, empirical research regarding self-talk interventions for sport and motor performance has burgeoned leading some to reflect on the field as "maturing" post-2011 [@hardyReflectionsMaturingResearch2018].

Whilst self-talk as a field may have matured in the post-2011 years with theoretical advancements in conceptualisation of the construct and mediators of its effects on performance, efforts to improve operationalisation/measurement, and efforts to improve methodology used in studying self-talk [@brinthauptSelftalkResearchChallenges2023; @geurtsMakingSenseSelf2018; @hardyReflectionsMaturingResearch2018; @latinjakSelfTalkInterdisciplinaryReview2023; @vanraalteSelftalkReviewSportspecific2016; @latinjakSpeakingClearly102019], it could be argued that understanding of the effectiveness of self-talk *interventions* (referred to in modern literature as 'strategic' self-talk; @latinjakSpeakingClearly102019) was "mature" prior to 2011.
The effect estimate from the meta-analysis of Hatzigeorgiadis et al. [-@hatzigeorgiadisSelfTalkSportsPerformance2011] might be considered by some to have been already fairly precise only spanning 0.2 SMD, and indeed aswell for many of the moderator estimates.
